---
title: 'Santa Barbara restaurant idle time'
author: "Jamie Afflerbach"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
  pdf_document:
    toc: true
---

### Setup

```{r setup, warning = F, message = F}

knitr::opts_chunk$set(warning = F, message = F)

#libraries
library(tidyverse)
library(mapview)
library(sf)
library(RColorBrewer)
library(lubridate)
```

### Load data

```{r data}
data <- read_csv("../data/master_data.csv") 
geocodes <- read_csv("../data/geocode_ref.csv")
```

### Cleaning data

Convert the `Opened` and `Died` columns into dates.

Important to find the addresses that didn't exist all the way back to 2007. Example - satellite, public market, blackbird. These were built more recently.

```{r}
df <- data %>%
  filter(nchar(Died) != 4) %>% #this removes 2005 and 2006. Those restaurants do not have a month attached
  left_join(geocodes) %>%
  mutate(Opened_date = dplyr::if_else(
            is.na(Opened), ymd("2007-01-01"), #if there is no open date, assume it opened Jan 1, 2007. We can adjust this later but for idle time since 2007 I Think this is fine.
            ymd(Opened, truncated = 1)),
         Closed_date = dplyr::if_else(
             is.na(Died), today(), #if it hasn't died, it's still open so give it the most current date.
             ymd(Died, truncated = 1)),
         individual_restaurant_length = Closed_date - Opened_date) %>% #total number of days restaurant was open
group_by(lat, lon) %>%
  mutate(total_days_open = sum(individual_restaurant_length), #calculate total days open
         earliest_open_date = min(Opened_date, na.rm = T),    #get earliest possible open date
         max_days_open = today() - earliest_open_date) %>%  #the maximum number of days the place could have a restaurant
  ungroup() %>%
  mutate(proportion_vacant = 100*(1-(as.numeric(total_days_open)/as.numeric(max_days_open)))) %>%
  select(Address, lon, lat, total_days_open, max_days_open, proportion_vacant) %>%
  distinct()
```


Next - find addresses with no open date since the most recent close date. This will help us find restaurants that maybe are no longer restaurants. For example Italian & Greek Market_636 State St, Santa Barbara
         
 

Ok some of these are negative probably because there are multiple locations at each address. Filtering just for positive

```{r}

idle_sf <- st_as_sf(df, coords = c("lon", "lat"), 
                 crs = 4326) %>%
  filter(proportion_vacant >= 0) %>%
  mutate(prop_vacant = as.integer(proportion_vacant))

pink2 = colorRampPalette(c('darkgreen', 'yellow', 'red'))

mapview(idle_sf, zcol = "prop_vacant", alpha = 0.8, layer.name = "% of time since 2007 vacant", col.region = pink2, at = seq(0, 100, 5))

```

